The Design of an Aliasing-Tolerant Parallel Programming Model

### 2012-11-09 doc/types-pldi2013/langdef.tex
trying to be consistent with math/text font choices

% start with read/write/reduce rules
\ruleat{\XA,0.0}{\XB,0.2}{T-Read}
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\forall i, reads(r_i) \in \Phi^*\end{array}}
{\typeenvx \text{read}(e_1) : T}

\ruleat{\XA,0.6}{\XB,0.8}{T-Write}
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\typeenvx e_2 : T \\
\forall i, writes(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})}

\ruleat{\XA,1.3}{\XB,1.5}{T-Reduce}
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \\
\typeenvx e_1 : T_1@(\oton{r}) \\
\typeenvx e_2 : T_2 \\
\forall i, reduces_{id}(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}

@@

\ruleat{\XA,2.5}{\XB,2.55}{T-UpRgn}
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})}

\ruleat{\XA,3.1}{\XB,3.05}{T-DnRgn}
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
%\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})}

@@

\ruleat{\XA,3.9}{\XB,4.1}{T-Color}
{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r) \\
\typeenvx e_2 : T@r \\
\typeenvx e_3 : int
\end{array}}
{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}

@@

\ruleatx{\XD,0.0}{\XE,0.2}{E-Read}{
\opsenvx e \mapsto l, E \andalso S' = \text{apply}(S, E) \\
v = \begin{cases} S'(l), & \text{if } l \not\in C \\
v' : H(l), & \text{otherwise} \end{cases}
}{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}

\ruleatx{\XD,0.9}{\XE,0.85}{E-Write}{
\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]}

\ruleatx{\XD,1.5}{\XE,1.45}{E-Reduce}{
\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}

@@

\ruleatx{\XD,4.7}{\XE,4.9}{E-Color}{
\opsenvx e_1 \mapsto K, E_1 \andalso S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \andalso S'' = \text{apply}(S', E_2) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
valid\_interleave(S, C, E', E_1, E_2, E_3)
}{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'}

\ruleatx{\XD,5.7}{\XE,5.55}{E-Partition}{
\opsenv{e_1}{K}{E_1} \andalso \rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \andalso S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'}

@@

\ruleatx{\XD,7.1}{\XE,6.95}{E-Unpack}{
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \andalso M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \andalso S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E'}

\ruleatx{\XA,8.1}{\XE,8.4}{E-Call}{
\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = \text{apply}(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = \text{apply}(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E}) \andalso \text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = \text{apply}(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})
}{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}

@@

\node[below right] at (\XA,9.15) {
$
\begin{array}{@{}l@{}}
\begin{array}{@{}l@{}l@{}} 
\begin{array}{@{}lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \\
& & \text{ where } S' = apply(S, E)
\end{array} &
\begin{array}{@{}ll}
mark\_coherence([], \hat Q) & = [] \\
mark\_coherence([ op(l, c, v) ] \dplus E, \hat Q) & = [ op(l, c', v) ] \dplus mark\_coherence(E, \hat Q), \\
\multicolumn{2}{@{}l}{ \hspace{1.0in} \text{ where } c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge simult(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge atomic(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases}}
\end{array}
\end{array}
\end{array}
$};

@@

\begin{figure}
\centering{
$\begin{array}{l}
\Omega \subseteq \Omega^* \\
r_i \leq r_j \in \Omega^*  \Rightarrow r_i \leq r_i \in \Omega^*\wedge r_j \leq r_j \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j \leq r_k \in \Omega^* \Rightarrow r_i \leq r_k \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j * r_k \in \Omega^* \Rightarrow r_i * r_k \in \Omega^* \\
r_i * r_j \in \Omega^* \Rightarrow r_j * r_i \in \Omega^* \\
\\
\Phi \subseteq \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reads(r_j) \in \Phi^* \Rightarrow reads(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge writes(r_j) \in \Phi^* \Rightarrow writes(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reduces_{id}(r_j) \in \Phi^* \Rightarrow reduces_{id}(r_i) \in \Phi^* \\
reads(r) \in \Phi^* \wedge writes(r) \in \Phi^* \Rightarrow reduces_{id}(r) \in \Phi^* \\
\ \ \ \mbox{for every function identifier $id$}
\end{array}$
}
\caption{Privilege and Constraint Closure}
\vspace{-5mm}
\label{fig:closure}
\end{figure}

@@

\texcomment{
\begin{figure*}
\centering{
$
\begin{array}{@{}l@{}}
\begin{array}{@{}l@{\hspace{-0.25in}}l@{}} 
\begin{array}{@{}lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \\
& & \text{ where } S' = apply(S, E)
\end{array} &
\begin{array}{@{}ll}
mark\_coh([], \hat Q) & = [] \\
mark\_coh([ op(l, c, v) ] \dplus E, \hat Q) & = [ op(l, c', v) ] \dplus mark\_coh(E, \hat Q), \\
\multicolumn{2}{@{}l}{ \hspace{1.0in} \text{ where } c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge simult(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge atomic(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases}}
\end{array}
\end{array} \\
\begin{array}{@{}lll}
any\_interleave([], [], \ldots, []) & = & true \\
any\_interleave([ \epsilon ] \dplus E', E_1, \ldots, [ \epsilon ] \dplus E_i, \ldots, E_n) & = & any\_interleave(E', E_1, \ldots, E_i, \ldots, E_n) \\
valid\_interleave(S, C, E', \oton{E}) & \Rightarrow & any\_interleave(E', \oton{E}) \\
\end{array}
\end{array}
$}
\caption{Operational Semantics Helper Functions \label{fig:opsemfns}}
\vspace{-5mm}
\end{figure*}
}

@@

Our first result shows a well-typed expression accesses the heap in ways consistent with its static privileges.
A judgment $\privconsist{E}{\Phi}$ holds if memory operations in memory trace $E$ have types and
locations covered by privileges $\Phi$:
\[ 
\begin{array}{l}
\privconsist{E}{\Phi} \Leftrightarrow \forall \epsilon \in E. \\
\quad\quad (\epsilon = read(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reads(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = write(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge writes(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reduces_{id}(r) \in \Phi)
\end{array}
\]

@@

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \Rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

$
\begin{array}{l@{}l}
\multicolumn{2}{l}{priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow} \\
& (priv_1 = reads \wedge priv_2 = reads)\ \vee \\
& (priv_1 = reduces_{id_1} \wedge priv_2 = reduces_{id_2} \wedge id_1 = id_2)\ \vee \\
& (r_1 * r_2)\ \vee \\
& (M(r_1) \cap M(r_2) = \emptyset) \\
\\
\multicolumn{2}{l}{\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2}
\end{array}$

>>>>

% start with read/write/reduce rules
\ruleat{\XA,0.0}{\XB,0.2}{T-Read}
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\forall i.\ \text{reads}(r_i) \in \Phi^*\end{array}}
{\typeenvx \text{read}(e_1) : T}

\ruleat{\XA,0.6}{\XB,0.8}{T-Write}
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\typeenvx e_2 : T \\
\forall i.\ \text{writes}(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})}

\ruleat{\XA,1.3}{\XB,1.5}{T-Reduce}
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \\
\typeenvx e_1 : T_1@(\oton{r}) \\
\typeenvx e_2 : T_2 \\
\forall i.\ \text{reduces}_{id}(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}

@@

\ruleat{\XA,2.5}{\XB,2.55}{T-UpRgn}
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenvx \text{upregion}(e,\oton{r}) : T@(\oton{r})}

\ruleat{\XA,3.1}{\XB,3.05}{T-DnRgn}
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
%\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenvx \text{downregion}(e,\oton{r}) : T@(\oton{r})}

@@

\ruleat{\XA,3.9}{\XB,4.1}{T-Color}
{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r) \\
\typeenvx e_2 : T@r \\
\typeenvx e_3 : \text{int}
\end{array}}
{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}

@@

\ruleatx{\XD,0.0}{\XE,0.2}{E-Read}{
\opsenvx e \mapsto l, E \andalso S' = apply(S, E) \\
v = \begin{cases} S'(l), & \text{if } l \not\in C \\
v' : H(l), & \text{otherwise} \end{cases}
}{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}

\ruleatx{\XD,0.9}{\XE,0.85}{E-Write}{
\opsenvx e_1 \mapsto l, E_1 \andalso S' = apply(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]}

\ruleatx{\XD,1.5}{\XE,1.45}{E-Reduce}{
\opsenvx e_1 \mapsto l, E_1 \andalso S' = apply(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}

@@

\ruleatx{\XD,4.7}{\XE,4.9}{E-Color}{
\opsenvx e_1 \mapsto K, E_1 \andalso S' = apply(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \andalso S'' = apply(S', E_2) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
valid\_interleave(S, C, E', E_1, E_2, E_3)
}{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'}

\ruleatx{\XD,5.7}{\XE,5.55}{E-Partition}{
\opsenv{e_1}{K}{E_1} \andalso \rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \andalso S' = apply(S, E_1) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'}

@@

\ruleatx{\XD,7.1}{\XE,6.95}{E-Unpack}{
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \andalso M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \andalso S' = apply(S, E_1) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)
}{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E'}

\ruleatx{\XA,8.1}{\XE,8.4}{E-Call}{
\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = apply(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = apply(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E}) \andalso \text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = apply(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})
}{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}

@@

\node[below right] at (\XA,9.15) {
$
\begin{array}{@{}l@{}}
\begin{array}{@{}l@{}l@{}} 
\begin{array}{@{}lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \\
& & \text{ where } S' = apply(S, E)
\end{array} &
\begin{array}{@{}ll}
mark\_coherence([], \hat Q) & = [] \\
mark\_coherence([ op(l, c, v) ] \dplus E, \hat Q) & = [ op(l, c', v) ] \dplus mark\_coherence(E, \hat Q), \\
\multicolumn{2}{@{}l}{ \hspace{1.0in} \text{ where } c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge \text{simult}(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge \text{atomic}(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases}}
\end{array}
\end{array}
\end{array}
$};

@@

\begin{figure}
\centering{
$\begin{array}{l}
\Omega \subseteq \Omega^* \\
r_i \leq r_j \in \Omega^*  \Rightarrow r_i \leq r_i \in \Omega^*\wedge r_j \leq r_j \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j \leq r_k \in \Omega^* \Rightarrow r_i \leq r_k \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j * r_k \in \Omega^* \Rightarrow r_i * r_k \in \Omega^* \\
r_i * r_j \in \Omega^* \Rightarrow r_j * r_i \in \Omega^* \\
\\
\Phi \subseteq \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge \text{reads}(r_j) \in \Phi^* \Rightarrow \text{reads}(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge \text{writes}(r_j) \in \Phi^* \Rightarrow \text{writes}(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge \text{reduces}_{id}(r_j) \in \Phi^* \Rightarrow \text{reduces}_{id}(r_i) \in \Phi^* \\
\text{reads}(r) \in \Phi^* \wedge \text{writes}(r) \in \Phi^* \Rightarrow \text{reduces}_{id}(r) \in \Phi^* \\
\ \ \ \mbox{for every function identifier $id$}
\end{array}$
}
\caption{Privilege and Constraint Closure}
\vspace{-5mm}
\label{fig:closure}
\end{figure}

@@

\texcomment{
\begin{figure*}
\centering{
$
\begin{array}{@{}l@{}}
\begin{array}{@{}l@{\hspace{-0.25in}}l@{}} 
\begin{array}{@{}lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \\
& & \text{ where } S' = apply(S, E)
\end{array} &
\begin{array}{@{}ll}
mark\_coh([], \hat Q) & = [] \\
mark\_coh([ op(l, c, v) ] \dplus E, \hat Q) & = [ op(l, c', v) ] \dplus mark\_coh(E, \hat Q), \\
\multicolumn{2}{@{}l}{ \hspace{1.0in} \text{ where } c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge \text{simult}(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge \text{atomic}(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases}}
\end{array}
\end{array} \\
\begin{array}{@{}lll}
any\_interleave([], [], \ldots, []) & = & true \\
any\_interleave([ \epsilon ] \dplus E', E_1, \ldots, [ \epsilon ] \dplus E_i, \ldots, E_n) & = & any\_interleave(E', E_1, \ldots, E_i, \ldots, E_n) \\
valid\_interleave(S, C, E', \oton{E}) & \Rightarrow & any\_interleave(E', \oton{E}) \\
\end{array}
\end{array}
$}
\caption{Operational Semantics Helper Functions \label{fig:opsemfns}}
\vspace{-5mm}
\end{figure*}
}

@@

Our first result shows a well-typed expression accesses the heap in ways consistent with its static privileges.
A judgment $\privconsist{E}{\Phi}$ holds if memory operations in memory trace $E$ have types and
locations covered by privileges $\Phi$:
\[ 
\begin{array}{l}
\privconsist{E}{\Phi} \Leftrightarrow \forall \epsilon \in E. \\
\quad\quad (\epsilon = read(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{reads}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = write(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{writes}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{reduces}_{id}(r) \in \Phi)
\end{array}
\]

@@

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\hat \Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\hat \Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

$
\begin{array}{l@{}l}
\multicolumn{2}{l}{priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow} \\
& (priv_1 = \text{reads} \wedge priv_2 = \text{reads})\ \vee \\
& (priv_1 = \text{reduces}_{id_1} \wedge priv_2 = \text{reduces}_{id_2} \wedge id_1 = id_2)\ \vee \\
& (r_1 * r_2)\ \vee \\
& (M(r_1) \cap M(r_2) = \emptyset) \\
\\
\multicolumn{2}{l}{\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2}
\end{array}$
